### 2023-03-30 tex/chapters/conclusion.tex
* Finished evaluation and conclusion for acm paper

\section{Conclusion}
\label{sec:conclusion}

In this work, we evaluated the prossibility of a novel verification approach called that can combine different dependent and independent type system in one verification framework. We have presented \textsc{LayeredTypes} -- A framework that realizes this verification approach on a simple functional language.

Using two distinct use-cases, we have shown how this approach can be used for various scenarios, including an exploration to the improvement of current liquid type implementations. Our implementation provides a scientific demonstrator for a layered and incremental verification approach. We believe this novel verification approach could exhibit great potential when building complex software systems. We hope that this work, alongside with our implementation, can provide valuable insights on the usefulness and potential challenges of such an approach.